### 2015-05-12 ipython_notebooks/hello.tex
Finally restarted working on the paper

% PyX 0.12.1 texdebug file
% mode: tex
% date: Tue May 12 13:40:03 2015
\scrollmode
\raiseerror%
\def\PyX{P\kern-.3em\lower.5ex\hbox{Y}\kern-.18em X}%
\gdef\PyXBoxHAlign{0}%
\newbox\PyXBox%
\newbox\PyXBoxHAligned%
\newdimen\PyXDimenHAlignLT%
\newdimen\PyXDimenHAlignRT%
\gdef\PyXFlushHAlign{0}%
\def\PyXragged{%
\leftskip=0pt plus \PyXFlushHAlign fil%
\rightskip=0pt plus 1fil%
\advance\rightskip0pt plus -\PyXFlushHAlign fil%
\parfillskip=0pt%
\pretolerance=9999%
\tolerance=9999%
\parindent=0pt%
\hyphenpenalty=9999%
\exhyphenpenalty=9999}%
\newbox\PyXBoxVBox%
\newdimen\PyXDimenVBox%
\newbox\PyXBoxVAlign%
\newdimen\PyXDimenVAlign%
\newdimen\PyXDimenVShift%
\long\def\ProcessPyXBox#1#2{%
\setbox\PyXBox=\hbox{{#1}}%
\PyXDimenHAlignLT=\PyXBoxHAlign\wd\PyXBox%
\PyXDimenHAlignRT=\wd\PyXBox%
\advance\PyXDimenHAlignRT by -\PyXDimenHAlignLT%
\gdef\PyXBoxHAlign{0}%
\immediate\write16{PyXBox:page=#2,lt=\the\PyXDimenHAlignLT,rt=\the\PyXDimenHAlignRT,ht=\the\ht\PyXBox,dp=\the\dp\PyXBox:}%
\setbox\PyXBoxHAligned=\hbox{\kern-\PyXDimenHAlignLT\box\PyXBox}%
\ht\PyXBoxHAligned0pt%
{\count0=80\count1=121\count2=88\count3=#2\shipout\box\PyXBoxHAligned}}%
\def\PyXInput#1{\immediate\write16{PyXInputMarker:executeid=#1:}}%
\def\PyXMarker#1{\hskip0pt\special{PyX:marker #1}}%%
\PyXInput{1}%
% This automatically generated file is part of PyX (http://pyx.sourceforge.net/).
% \latexstylename="10pt"
% \latexclassname="article"
% \latexclassopt="10pt"
% \latexinit=""
\def\tiny{\font\pyxfont=cmr5\pyxfont\font\pyxfonttfa=cmr5\textfont0=\pyxfonttfa\font\pyxfontsfa=cmr5\scriptfont0=\pyxfontsfa\font\pyxfontssfa=cmr5\scriptscriptfont0=\pyxfontssfa\font\pyxfonttfb=cmmi5\textfont1=\pyxfonttfb\font\pyxfontsfb=cmmi5\scriptfont1=\pyxfontsfb\font\pyxfontssfb=cmmi5\scriptscriptfont1=\pyxfontssfb\font\pyxfonttfc=cmsy5\textfont2=\pyxfonttfc\font\pyxfontsfc=cmsy5\scriptfont2=\pyxfontsfc\font\pyxfontssfc=cmsy5\scriptscriptfont2=\pyxfontssfc\font\pyxfonttfd=cmex10\textfont3=\pyxfonttfd\font\pyxfontsfd=cmex10\scriptfont3=\pyxfontsfd\font\pyxfontssfd=cmex10\scriptscriptfont3=\pyxfontssfd}
\def\scriptsize{\font\pyxfont=cmr7\pyxfont\font\pyxfonttfa=cmr7\textfont0=\pyxfonttfa\font\pyxfontsfa=cmr5\scriptfont0=\pyxfontsfa\font\pyxfontssfa=cmr5\scriptscriptfont0=\pyxfontssfa\font\pyxfonttfb=cmmi7\textfont1=\pyxfonttfb\font\pyxfontsfb=cmmi5\scriptfont1=\pyxfontsfb\font\pyxfontssfb=cmmi5\scriptscriptfont1=\pyxfontssfb\font\pyxfonttfc=cmsy7\textfont2=\pyxfonttfc\font\pyxfontsfc=cmsy5\scriptfont2=\pyxfontsfc\font\pyxfontssfc=cmsy5\scriptscriptfont2=\pyxfontssfc\font\pyxfonttfd=cmex10\textfont3=\pyxfonttfd\font\pyxfontsfd=cmex10\scriptfont3=\pyxfontsfd\font\pyxfontssfd=cmex10\scriptscriptfont3=\pyxfontssfd}
\def\footnotesize{\font\pyxfont=cmr8\pyxfont\font\pyxfonttfa=cmr8\textfont0=\pyxfonttfa\font\pyxfontsfa=cmr6\scriptfont0=\pyxfontsfa\font\pyxfontssfa=cmr5\scriptscriptfont0=\pyxfontssfa\font\pyxfonttfb=cmmi8\textfont1=\pyxfonttfb\font\pyxfontsfb=cmmi6\scriptfont1=\pyxfontsfb\font\pyxfontssfb=cmmi5\scriptscriptfont1=\pyxfontssfb\font\pyxfonttfc=cmsy8\textfont2=\pyxfonttfc\font\pyxfontsfc=cmsy6\scriptfont2=\pyxfontsfc\font\pyxfontssfc=cmsy5\scriptscriptfont2=\pyxfontssfc\font\pyxfonttfd=cmex10\textfont3=\pyxfonttfd\font\pyxfontsfd=cmex10\scriptfont3=\pyxfontsfd\font\pyxfontssfd=cmex10\scriptscriptfont3=\pyxfontssfd}
\def\small{\font\pyxfont=cmr9\pyxfont\font\pyxfonttfa=cmr9\textfont0=\pyxfonttfa\font\pyxfontsfa=cmr6\scriptfont0=\pyxfontsfa\font\pyxfontssfa=cmr5\scriptscriptfont0=\pyxfontssfa\font\pyxfonttfb=cmmi9\textfont1=\pyxfonttfb\font\pyxfontsfb=cmmi6\scriptfont1=\pyxfontsfb\font\pyxfontssfb=cmmi5\scriptscriptfont1=\pyxfontssfb\font\pyxfonttfc=cmsy9\textfont2=\pyxfonttfc\font\pyxfontsfc=cmsy6\scriptfont2=\pyxfontsfc\font\pyxfontssfc=cmsy5\scriptscriptfont2=\pyxfontssfc\font\pyxfonttfd=cmex10\textfont3=\pyxfonttfd\font\pyxfontsfd=cmex10\scriptfont3=\pyxfontsfd\font\pyxfontssfd=cmex10\scriptscriptfont3=\pyxfontssfd}
\def\normalsize{\font\pyxfont=cmr10\pyxfont\font\pyxfonttfa=cmr10\textfont0=\pyxfonttfa\font\pyxfontsfa=cmr7\scriptfont0=\pyxfontsfa\font\pyxfontssfa=cmr5\scriptscriptfont0=\pyxfontssfa\font\pyxfonttfb=cmmi10\textfont1=\pyxfonttfb\font\pyxfontsfb=cmmi7\scriptfont1=\pyxfontsfb\font\pyxfontssfb=cmmi5\scriptscriptfont1=\pyxfontssfb\font\pyxfonttfc=cmsy10\textfont2=\pyxfonttfc\font\pyxfontsfc=cmsy7\scriptfont2=\pyxfontsfc\font\pyxfontssfc=cmsy5\scriptscriptfont2=\pyxfontssfc\font\pyxfonttfd=cmex10\textfont3=\pyxfonttfd\font\pyxfontsfd=cmex10\scriptfont3=\pyxfontsfd\font\pyxfontssfd=cmex10\scriptscriptfont3=\pyxfontssfd}
\def\large{\font\pyxfont=cmr12\pyxfont\font\pyxfonttfa=cmr12\textfont0=\pyxfonttfa\font\pyxfontsfa=cmr8\scriptfont0=\pyxfontsfa\font\pyxfontssfa=cmr6\scriptscriptfont0=\pyxfontssfa\font\pyxfonttfb=cmmi12\textfont1=\pyxfonttfb\font\pyxfontsfb=cmmi8\scriptfont1=\pyxfontsfb\font\pyxfontssfb=cmmi6\scriptscriptfont1=\pyxfontssfb\font\pyxfonttfc=cmsy10 at 12.0pt\textfont2=\pyxfonttfc\font\pyxfontsfc=cmsy8\scriptfont2=\pyxfontsfc\font\pyxfontssfc=cmsy6\scriptscriptfont2=\pyxfontssfc\font\pyxfonttfd=cmex10\textfont3=\pyxfonttfd\font\pyxfontsfd=cmex10\scriptfont3=\pyxfontsfd\font\pyxfontssfd=cmex10\scriptscriptfont3=\pyxfontssfd}
\def\Large{\font\pyxfont=cmr12 at 14.4pt\pyxfont\font\pyxfonttfa=cmr12 at 14.4pt\textfont0=\pyxfonttfa\font\pyxfontsfa=cmr10\scriptfont0=\pyxfontsfa\font\pyxfontssfa=cmr7\scriptscriptfont0=\pyxfontssfa\font\pyxfonttfb=cmmi12 at 14.4pt\textfont1=\pyxfonttfb\font\pyxfontsfb=cmmi10\scriptfont1=\pyxfontsfb\font\pyxfontssfb=cmmi7\scriptscriptfont1=\pyxfontssfb\font\pyxfonttfc=cmsy10 at 14.4pt\textfont2=\pyxfonttfc\font\pyxfontsfc=cmsy10\scriptfont2=\pyxfontsfc\font\pyxfontssfc=cmsy7\scriptscriptfont2=\pyxfontssfc\font\pyxfonttfd=cmex10\textfont3=\pyxfonttfd\font\pyxfontsfd=cmex10\scriptfont3=\pyxfontsfd\font\pyxfontssfd=cmex10\scriptscriptfont3=\pyxfontssfd}
\def\LARGE{\font\pyxfont=cmr17\pyxfont\font\pyxfonttfa=cmr17\textfont0=\pyxfonttfa\font\pyxfontsfa=cmr12\scriptfont0=\pyxfontsfa\font\pyxfontssfa=cmr10\scriptscriptfont0=\pyxfontssfa\font\pyxfonttfb=cmmi12 at 17.28pt\textfont1=\pyxfonttfb\font\pyxfontsfb=cmmi12\scriptfont1=\pyxfontsfb\font\pyxfontssfb=cmmi10\scriptscriptfont1=\pyxfontssfb\font\pyxfonttfc=cmsy10 at 17.28pt\textfont2=\pyxfonttfc\font\pyxfontsfc=cmsy10 at 12.0pt\scriptfont2=\pyxfontsfc\font\pyxfontssfc=cmsy10\scriptscriptfont2=\pyxfontssfc\font\pyxfonttfd=cmex10\textfont3=\pyxfonttfd\font\pyxfontsfd=cmex10\scriptfont3=\pyxfontsfd\font\pyxfontssfd=cmex10\scriptscriptfont3=\pyxfontssfd}
\def\huge{\font\pyxfont=cmr17 at 20.74pt\pyxfont\font\pyxfonttfa=cmr17 at 20.74pt\textfont0=\pyxfonttfa\font\pyxfontsfa=cmr12 at 14.4pt\scriptfont0=\pyxfontsfa\font\pyxfontssfa=cmr12\scriptscriptfont0=\pyxfontssfa\font\pyxfonttfb=cmmi12 at 20.74pt\textfont1=\pyxfonttfb\font\pyxfontsfb=cmmi12 at 14.4pt\scriptfont1=\pyxfontsfb\font\pyxfontssfb=cmmi12\scriptscriptfont1=\pyxfontssfb\font\pyxfonttfc=cmsy10 at 20.74pt\textfont2=\pyxfonttfc\font\pyxfontsfc=cmsy10 at 14.4pt\scriptfont2=\pyxfontsfc\font\pyxfontssfc=cmsy10 at 12.0pt\scriptscriptfont2=\pyxfontssfc\font\pyxfonttfd=cmex10\textfont3=\pyxfonttfd\font\pyxfontsfd=cmex10\scriptfont3=\pyxfontsfd\font\pyxfontssfd=cmex10\scriptscriptfont3=\pyxfontssfd}
\def\Huge{\font\pyxfont=cmr17 at 24.88pt\pyxfont\font\pyxfonttfa=cmr17 at 24.88pt\textfont0=\pyxfonttfa\font\pyxfontsfa=cmr17 at 20.74pt\scriptfont0=\pyxfontsfa\font\pyxfontssfa=cmr17\scriptscriptfont0=\pyxfontssfa\font\pyxfonttfb=cmmi12 at 24.88pt\textfont1=\pyxfonttfb\font\pyxfontsfb=cmmi12 at 20.74pt\scriptfont1=\pyxfontsfb\font\pyxfontssfb=cmmi12 at 17.28pt\scriptscriptfont1=\pyxfontssfb\font\pyxfonttfc=cmsy10 at 24.88pt\textfont2=\pyxfonttfc\font\pyxfontsfc=cmsy10 at 20.74pt\scriptfont2=\pyxfontsfc\font\pyxfontssfc=cmsy10 at 17.28pt\scriptscriptfont2=\pyxfontssfc\font\pyxfonttfd=cmex10\textfont3=\pyxfonttfd\font\pyxfontsfd=cmex10\scriptfont3=\pyxfontsfd\font\pyxfontssfd=cmex10\scriptscriptfont3=\pyxfontssfd}
%
\PyXInput{2}%
\normalsize%
%
\PyXInput{3}%
\newdimen\linewidth\newdimen\textwidth%
%
\PyXInput{4}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}$\displaystyle{1}$}}\lower\PyXDimenVShift\box0%
}{1}%
\PyXInput{5}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}$\displaystyle{2}$}}\lower\PyXDimenVShift\box0%
}{2}%
\PyXInput{6}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}$\displaystyle{5}$}}\lower\PyXDimenVShift\box0%
}{3}%
\PyXInput{7}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}$\displaystyle{10}$}}\lower\PyXDimenVShift\box0%
}{4}%
\PyXInput{8}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}output power $\times10^{10}$ W/Hz}}\lower\PyXDimenVShift\box0%
}{5}%
\PyXInput{9}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}$\displaystyle{0.1}$}}\lower\PyXDimenVShift\box0%
}{6}%
\PyXInput{10}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}$\displaystyle{1}$}}\lower\PyXDimenVShift\box0%
}{7}%
\PyXInput{11}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}$\displaystyle{10}$}}\lower\PyXDimenVShift\box0%
}{8}%
\PyXInput{12}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}$\displaystyle{100}$}}\lower\PyXDimenVShift\box0%
}{9}%
\PyXInput{13}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}$\displaystyle{1000}$}}\lower\PyXDimenVShift\box0%
}{10}%
\PyXInput{14}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}output power $\times10^{10}$ W/Hz}}\lower\PyXDimenVShift\box0%
}{11}%
\PyXInput{15}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}$\displaystyle{{0}}$}}\lower\PyXDimenVShift\box0%
}{12}%
\PyXInput{16}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}$\displaystyle{{{2.5}\cdot10^{3}}}$}}\lower\PyXDimenVShift\box0%
}{13}%
\PyXInput{17}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}$\displaystyle{{{5}\cdot10^{3}}}$}}\lower\PyXDimenVShift\box0%
}{14}%
\PyXInput{18}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}$\displaystyle{{{7.5}\cdot10^{3}}}$}}\lower\PyXDimenVShift\box0%
}{15}%
\PyXInput{19}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}$\displaystyle{{{1}\cdot10^{4}}}$}}\lower\PyXDimenVShift\box0%
}{16}%
\PyXInput{20}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}active Q}}\lower\PyXDimenVShift\box0%
}{17}%
\PyXInput{21}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}$\displaystyle{{0}}$}}\lower\PyXDimenVShift\box0%
}{18}%
\PyXInput{22}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}$\displaystyle{{{2}\cdot10^{3}}}$}}\lower\PyXDimenVShift\box0%
}{19}%
\PyXInput{23}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}$\displaystyle{{{4}\cdot10^{3}}}$}}\lower\PyXDimenVShift\box0%
}{20}%
\PyXInput{24}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}$\displaystyle{{{6}\cdot10^{3}}}$}}\lower\PyXDimenVShift\box0%
}{21}%
\PyXInput{25}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}$\displaystyle{{{8}\cdot10^{3}}}$}}\lower\PyXDimenVShift\box0%
}{22}%
\PyXInput{26}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}$\displaystyle{{{1}\cdot10^{4}}}$}}\lower\PyXDimenVShift\box0%
}{23}%
\PyXInput{27}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}active Q}}\lower\PyXDimenVShift\box0%
}{24}%
\PyXInput{28}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{signal}}\lower\PyXDimenVShift\box0%
}{25}%
\PyXInput{29}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{noise}}\lower\PyXDimenVShift\box0%
}{26}%
\PyXInput{30}%
\end%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% preparing restart of tex
% PyX 0.12.1 texdebug file
% mode: tex
% date: Tue May 12 13:42:56 2015
\scrollmode
\raiseerror%
\def\PyX{P\kern-.3em\lower.5ex\hbox{Y}\kern-.18em X}%
\gdef\PyXBoxHAlign{0}%
\newbox\PyXBox%
\newbox\PyXBoxHAligned%
\newdimen\PyXDimenHAlignLT%
\newdimen\PyXDimenHAlignRT%
\gdef\PyXFlushHAlign{0}%
\def\PyXragged{%
\leftskip=0pt plus \PyXFlushHAlign fil%
\rightskip=0pt plus 1fil%
\advance\rightskip0pt plus -\PyXFlushHAlign fil%
\parfillskip=0pt%
\pretolerance=9999%
\tolerance=9999%
\parindent=0pt%
\hyphenpenalty=9999%
\exhyphenpenalty=9999}%
\newbox\PyXBoxVBox%
\newdimen\PyXDimenVBox%
\newbox\PyXBoxVAlign%
\newdimen\PyXDimenVAlign%
\newdimen\PyXDimenVShift%
\long\def\ProcessPyXBox#1#2{%
\setbox\PyXBox=\hbox{{#1}}%
\PyXDimenHAlignLT=\PyXBoxHAlign\wd\PyXBox%
\PyXDimenHAlignRT=\wd\PyXBox%
\advance\PyXDimenHAlignRT by -\PyXDimenHAlignLT%
\gdef\PyXBoxHAlign{0}%
\immediate\write16{PyXBox:page=#2,lt=\the\PyXDimenHAlignLT,rt=\the\PyXDimenHAlignRT,ht=\the\ht\PyXBox,dp=\the\dp\PyXBox:}%
\setbox\PyXBoxHAligned=\hbox{\kern-\PyXDimenHAlignLT\box\PyXBox}%
\ht\PyXBoxHAligned0pt%
{\count0=80\count1=121\count2=88\count3=#2\shipout\box\PyXBoxHAligned}}%
\def\PyXInput#1{\immediate\write16{PyXInputMarker:executeid=#1:}}%
\def\PyXMarker#1{\hskip0pt\special{PyX:marker #1}}%%
\PyXInput{1}%
% This automatically generated file is part of PyX (http://pyx.sourceforge.net/).
% \latexstylename="10pt"
% \latexclassname="article"
% \latexclassopt="10pt"
% \latexinit=""
\def\tiny{\font\pyxfont=cmr5\pyxfont\font\pyxfonttfa=cmr5\textfont0=\pyxfonttfa\font\pyxfontsfa=cmr5\scriptfont0=\pyxfontsfa\font\pyxfontssfa=cmr5\scriptscriptfont0=\pyxfontssfa\font\pyxfonttfb=cmmi5\textfont1=\pyxfonttfb\font\pyxfontsfb=cmmi5\scriptfont1=\pyxfontsfb\font\pyxfontssfb=cmmi5\scriptscriptfont1=\pyxfontssfb\font\pyxfonttfc=cmsy5\textfont2=\pyxfonttfc\font\pyxfontsfc=cmsy5\scriptfont2=\pyxfontsfc\font\pyxfontssfc=cmsy5\scriptscriptfont2=\pyxfontssfc\font\pyxfonttfd=cmex10\textfont3=\pyxfonttfd\font\pyxfontsfd=cmex10\scriptfont3=\pyxfontsfd\font\pyxfontssfd=cmex10\scriptscriptfont3=\pyxfontssfd}
\def\scriptsize{\font\pyxfont=cmr7\pyxfont\font\pyxfonttfa=cmr7\textfont0=\pyxfonttfa\font\pyxfontsfa=cmr5\scriptfont0=\pyxfontsfa\font\pyxfontssfa=cmr5\scriptscriptfont0=\pyxfontssfa\font\pyxfonttfb=cmmi7\textfont1=\pyxfonttfb\font\pyxfontsfb=cmmi5\scriptfont1=\pyxfontsfb\font\pyxfontssfb=cmmi5\scriptscriptfont1=\pyxfontssfb\font\pyxfonttfc=cmsy7\textfont2=\pyxfonttfc\font\pyxfontsfc=cmsy5\scriptfont2=\pyxfontsfc\font\pyxfontssfc=cmsy5\scriptscriptfont2=\pyxfontssfc\font\pyxfonttfd=cmex10\textfont3=\pyxfonttfd\font\pyxfontsfd=cmex10\scriptfont3=\pyxfontsfd\font\pyxfontssfd=cmex10\scriptscriptfont3=\pyxfontssfd}
\def\footnotesize{\font\pyxfont=cmr8\pyxfont\font\pyxfonttfa=cmr8\textfont0=\pyxfonttfa\font\pyxfontsfa=cmr6\scriptfont0=\pyxfontsfa\font\pyxfontssfa=cmr5\scriptscriptfont0=\pyxfontssfa\font\pyxfonttfb=cmmi8\textfont1=\pyxfonttfb\font\pyxfontsfb=cmmi6\scriptfont1=\pyxfontsfb\font\pyxfontssfb=cmmi5\scriptscriptfont1=\pyxfontssfb\font\pyxfonttfc=cmsy8\textfont2=\pyxfonttfc\font\pyxfontsfc=cmsy6\scriptfont2=\pyxfontsfc\font\pyxfontssfc=cmsy5\scriptscriptfont2=\pyxfontssfc\font\pyxfonttfd=cmex10\textfont3=\pyxfonttfd\font\pyxfontsfd=cmex10\scriptfont3=\pyxfontsfd\font\pyxfontssfd=cmex10\scriptscriptfont3=\pyxfontssfd}
\def\small{\font\pyxfont=cmr9\pyxfont\font\pyxfonttfa=cmr9\textfont0=\pyxfonttfa\font\pyxfontsfa=cmr6\scriptfont0=\pyxfontsfa\font\pyxfontssfa=cmr5\scriptscriptfont0=\pyxfontssfa\font\pyxfonttfb=cmmi9\textfont1=\pyxfonttfb\font\pyxfontsfb=cmmi6\scriptfont1=\pyxfontsfb\font\pyxfontssfb=cmmi5\scriptscriptfont1=\pyxfontssfb\font\pyxfonttfc=cmsy9\textfont2=\pyxfonttfc\font\pyxfontsfc=cmsy6\scriptfont2=\pyxfontsfc\font\pyxfontssfc=cmsy5\scriptscriptfont2=\pyxfontssfc\font\pyxfonttfd=cmex10\textfont3=\pyxfonttfd\font\pyxfontsfd=cmex10\scriptfont3=\pyxfontsfd\font\pyxfontssfd=cmex10\scriptscriptfont3=\pyxfontssfd}
\def\normalsize{\font\pyxfont=cmr10\pyxfont\font\pyxfonttfa=cmr10\textfont0=\pyxfonttfa\font\pyxfontsfa=cmr7\scriptfont0=\pyxfontsfa\font\pyxfontssfa=cmr5\scriptscriptfont0=\pyxfontssfa\font\pyxfonttfb=cmmi10\textfont1=\pyxfonttfb\font\pyxfontsfb=cmmi7\scriptfont1=\pyxfontsfb\font\pyxfontssfb=cmmi5\scriptscriptfont1=\pyxfontssfb\font\pyxfonttfc=cmsy10\textfont2=\pyxfonttfc\font\pyxfontsfc=cmsy7\scriptfont2=\pyxfontsfc\font\pyxfontssfc=cmsy5\scriptscriptfont2=\pyxfontssfc\font\pyxfonttfd=cmex10\textfont3=\pyxfonttfd\font\pyxfontsfd=cmex10\scriptfont3=\pyxfontsfd\font\pyxfontssfd=cmex10\scriptscriptfont3=\pyxfontssfd}
\def\large{\font\pyxfont=cmr12\pyxfont\font\pyxfonttfa=cmr12\textfont0=\pyxfonttfa\font\pyxfontsfa=cmr8\scriptfont0=\pyxfontsfa\font\pyxfontssfa=cmr6\scriptscriptfont0=\pyxfontssfa\font\pyxfonttfb=cmmi12\textfont1=\pyxfonttfb\font\pyxfontsfb=cmmi8\scriptfont1=\pyxfontsfb\font\pyxfontssfb=cmmi6\scriptscriptfont1=\pyxfontssfb\font\pyxfonttfc=cmsy10 at 12.0pt\textfont2=\pyxfonttfc\font\pyxfontsfc=cmsy8\scriptfont2=\pyxfontsfc\font\pyxfontssfc=cmsy6\scriptscriptfont2=\pyxfontssfc\font\pyxfonttfd=cmex10\textfont3=\pyxfonttfd\font\pyxfontsfd=cmex10\scriptfont3=\pyxfontsfd\font\pyxfontssfd=cmex10\scriptscriptfont3=\pyxfontssfd}
\def\Large{\font\pyxfont=cmr12 at 14.4pt\pyxfont\font\pyxfonttfa=cmr12 at 14.4pt\textfont0=\pyxfonttfa\font\pyxfontsfa=cmr10\scriptfont0=\pyxfontsfa\font\pyxfontssfa=cmr7\scriptscriptfont0=\pyxfontssfa\font\pyxfonttfb=cmmi12 at 14.4pt\textfont1=\pyxfonttfb\font\pyxfontsfb=cmmi10\scriptfont1=\pyxfontsfb\font\pyxfontssfb=cmmi7\scriptscriptfont1=\pyxfontssfb\font\pyxfonttfc=cmsy10 at 14.4pt\textfont2=\pyxfonttfc\font\pyxfontsfc=cmsy10\scriptfont2=\pyxfontsfc\font\pyxfontssfc=cmsy7\scriptscriptfont2=\pyxfontssfc\font\pyxfonttfd=cmex10\textfont3=\pyxfonttfd\font\pyxfontsfd=cmex10\scriptfont3=\pyxfontsfd\font\pyxfontssfd=cmex10\scriptscriptfont3=\pyxfontssfd}
\def\LARGE{\font\pyxfont=cmr17\pyxfont\font\pyxfonttfa=cmr17\textfont0=\pyxfonttfa\font\pyxfontsfa=cmr12\scriptfont0=\pyxfontsfa\font\pyxfontssfa=cmr10\scriptscriptfont0=\pyxfontssfa\font\pyxfonttfb=cmmi12 at 17.28pt\textfont1=\pyxfonttfb\font\pyxfontsfb=cmmi12\scriptfont1=\pyxfontsfb\font\pyxfontssfb=cmmi10\scriptscriptfont1=\pyxfontssfb\font\pyxfonttfc=cmsy10 at 17.28pt\textfont2=\pyxfonttfc\font\pyxfontsfc=cmsy10 at 12.0pt\scriptfont2=\pyxfontsfc\font\pyxfontssfc=cmsy10\scriptscriptfont2=\pyxfontssfc\font\pyxfonttfd=cmex10\textfont3=\pyxfonttfd\font\pyxfontsfd=cmex10\scriptfont3=\pyxfontsfd\font\pyxfontssfd=cmex10\scriptscriptfont3=\pyxfontssfd}
\def\huge{\font\pyxfont=cmr17 at 20.74pt\pyxfont\font\pyxfonttfa=cmr17 at 20.74pt\textfont0=\pyxfonttfa\font\pyxfontsfa=cmr12 at 14.4pt\scriptfont0=\pyxfontsfa\font\pyxfontssfa=cmr12\scriptscriptfont0=\pyxfontssfa\font\pyxfonttfb=cmmi12 at 20.74pt\textfont1=\pyxfonttfb\font\pyxfontsfb=cmmi12 at 14.4pt\scriptfont1=\pyxfontsfb\font\pyxfontssfb=cmmi12\scriptscriptfont1=\pyxfontssfb\font\pyxfonttfc=cmsy10 at 20.74pt\textfont2=\pyxfonttfc\font\pyxfontsfc=cmsy10 at 14.4pt\scriptfont2=\pyxfontsfc\font\pyxfontssfc=cmsy10 at 12.0pt\scriptscriptfont2=\pyxfontssfc\font\pyxfonttfd=cmex10\textfont3=\pyxfonttfd\font\pyxfontsfd=cmex10\scriptfont3=\pyxfontsfd\font\pyxfontssfd=cmex10\scriptscriptfont3=\pyxfontssfd}
\def\Huge{\font\pyxfont=cmr17 at 24.88pt\pyxfont\font\pyxfonttfa=cmr17 at 24.88pt\textfont0=\pyxfonttfa\font\pyxfontsfa=cmr17 at 20.74pt\scriptfont0=\pyxfontsfa\font\pyxfontssfa=cmr17\scriptscriptfont0=\pyxfontssfa\font\pyxfonttfb=cmmi12 at 24.88pt\textfont1=\pyxfonttfb\font\pyxfontsfb=cmmi12 at 20.74pt\scriptfont1=\pyxfontsfb\font\pyxfontssfb=cmmi12 at 17.28pt\scriptscriptfont1=\pyxfontssfb\font\pyxfonttfc=cmsy10 at 24.88pt\textfont2=\pyxfonttfc\font\pyxfontsfc=cmsy10 at 20.74pt\scriptfont2=\pyxfontsfc\font\pyxfontssfc=cmsy10 at 17.28pt\scriptscriptfont2=\pyxfontssfc\font\pyxfonttfd=cmex10\textfont3=\pyxfonttfd\font\pyxfontsfd=cmex10\scriptfont3=\pyxfontsfd\font\pyxfontssfd=cmex10\scriptscriptfont3=\pyxfontssfd}
%
\PyXInput{2}%
\normalsize%
%
\PyXInput{3}%
\newdimen\linewidth\newdimen\textwidth%
%
\PyXInput{4}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}$\displaystyle{0.1}$}}\lower\PyXDimenVShift\box0%
}{1}%
\PyXInput{5}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}$\displaystyle{1}$}}\lower\PyXDimenVShift\box0%
}{2}%
\PyXInput{6}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}$\displaystyle{10}$}}\lower\PyXDimenVShift\box0%
}{3}%
\PyXInput{7}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}$\displaystyle{100}$}}\lower\PyXDimenVShift\box0%
}{4}%
\PyXInput{8}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}$\displaystyle{1000}$}}\lower\PyXDimenVShift\box0%
}{5}%
\PyXInput{9}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}output power ($\times10^{10}$) W/Hz}}\lower\PyXDimenVShift\box0%
}{6}%
\PyXInput{10}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}$\displaystyle{{0}}$}}\lower\PyXDimenVShift\box0%
}{7}%
\PyXInput{11}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}$\displaystyle{{{2.5}\cdot10^{3}}}$}}\lower\PyXDimenVShift\box0%
}{8}%
\PyXInput{12}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}$\displaystyle{{{5}\cdot10^{3}}}$}}\lower\PyXDimenVShift\box0%
}{9}%
\PyXInput{13}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}$\displaystyle{{{7.5}\cdot10^{3}}}$}}\lower\PyXDimenVShift\box0%
}{10}%
\PyXInput{14}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}$\displaystyle{{{1}\cdot10^{4}}}$}}\lower\PyXDimenVShift\box0%
}{11}%
\PyXInput{15}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}active Q}}\lower\PyXDimenVShift\box0%
}{12}%
\PyXInput{16}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}$\displaystyle{{0}}$}}\lower\PyXDimenVShift\box0%
}{13}%
\PyXInput{17}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}$\displaystyle{{{2}\cdot10^{3}}}$}}\lower\PyXDimenVShift\box0%
}{14}%
\PyXInput{18}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}$\displaystyle{{{4}\cdot10^{3}}}$}}\lower\PyXDimenVShift\box0%
}{15}%
\PyXInput{19}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}$\displaystyle{{{6}\cdot10^{3}}}$}}\lower\PyXDimenVShift\box0%
}{16}%
\PyXInput{20}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}$\displaystyle{{{8}\cdot10^{3}}}$}}\lower\PyXDimenVShift\box0%
}{17}%
\PyXInput{21}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}$\displaystyle{{{1}\cdot10^{4}}}$}}\lower\PyXDimenVShift\box0%
}{18}%
\PyXInput{22}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}active Q}}\lower\PyXDimenVShift\box0%
}{19}%
\PyXInput{23}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{signal}}\lower\PyXDimenVShift\box0%
}{20}%
\PyXInput{24}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{noise}}\lower\PyXDimenVShift\box0%
}{21}%
\PyXInput{25}%
\end%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% preparing restart of tex
% PyX 0.12.1 texdebug file
% mode: tex
% date: Tue May 12 13:43:35 2015
\scrollmode
\raiseerror%
\def\PyX{P\kern-.3em\lower.5ex\hbox{Y}\kern-.18em X}%
\gdef\PyXBoxHAlign{0}%
\newbox\PyXBox%
\newbox\PyXBoxHAligned%
\newdimen\PyXDimenHAlignLT%
\newdimen\PyXDimenHAlignRT%
\gdef\PyXFlushHAlign{0}%
\def\PyXragged{%
\leftskip=0pt plus \PyXFlushHAlign fil%
\rightskip=0pt plus 1fil%
\advance\rightskip0pt plus -\PyXFlushHAlign fil%
\parfillskip=0pt%
\pretolerance=9999%
\tolerance=9999%
\parindent=0pt%
\hyphenpenalty=9999%
\exhyphenpenalty=9999}%
\newbox\PyXBoxVBox%
\newdimen\PyXDimenVBox%
\newbox\PyXBoxVAlign%
\newdimen\PyXDimenVAlign%
\newdimen\PyXDimenVShift%
\long\def\ProcessPyXBox#1#2{%
\setbox\PyXBox=\hbox{{#1}}%
\PyXDimenHAlignLT=\PyXBoxHAlign\wd\PyXBox%
\PyXDimenHAlignRT=\wd\PyXBox%
\advance\PyXDimenHAlignRT by -\PyXDimenHAlignLT%
\gdef\PyXBoxHAlign{0}%
\immediate\write16{PyXBox:page=#2,lt=\the\PyXDimenHAlignLT,rt=\the\PyXDimenHAlignRT,ht=\the\ht\PyXBox,dp=\the\dp\PyXBox:}%
\setbox\PyXBoxHAligned=\hbox{\kern-\PyXDimenHAlignLT\box\PyXBox}%
\ht\PyXBoxHAligned0pt%
{\count0=80\count1=121\count2=88\count3=#2\shipout\box\PyXBoxHAligned}}%
\def\PyXInput#1{\immediate\write16{PyXInputMarker:executeid=#1:}}%
\def\PyXMarker#1{\hskip0pt\special{PyX:marker #1}}%%
\PyXInput{1}%
% This automatically generated file is part of PyX (http://pyx.sourceforge.net/).
% \latexstylename="10pt"
% \latexclassname="article"
% \latexclassopt="10pt"
% \latexinit=""
\def\tiny{\font\pyxfont=cmr5\pyxfont\font\pyxfonttfa=cmr5\textfont0=\pyxfonttfa\font\pyxfontsfa=cmr5\scriptfont0=\pyxfontsfa\font\pyxfontssfa=cmr5\scriptscriptfont0=\pyxfontssfa\font\pyxfonttfb=cmmi5\textfont1=\pyxfonttfb\font\pyxfontsfb=cmmi5\scriptfont1=\pyxfontsfb\font\pyxfontssfb=cmmi5\scriptscriptfont1=\pyxfontssfb\font\pyxfonttfc=cmsy5\textfont2=\pyxfonttfc\font\pyxfontsfc=cmsy5\scriptfont2=\pyxfontsfc\font\pyxfontssfc=cmsy5\scriptscriptfont2=\pyxfontssfc\font\pyxfonttfd=cmex10\textfont3=\pyxfonttfd\font\pyxfontsfd=cmex10\scriptfont3=\pyxfontsfd\font\pyxfontssfd=cmex10\scriptscriptfont3=\pyxfontssfd}
\def\scriptsize{\font\pyxfont=cmr7\pyxfont\font\pyxfonttfa=cmr7\textfont0=\pyxfonttfa\font\pyxfontsfa=cmr5\scriptfont0=\pyxfontsfa\font\pyxfontssfa=cmr5\scriptscriptfont0=\pyxfontssfa\font\pyxfonttfb=cmmi7\textfont1=\pyxfonttfb\font\pyxfontsfb=cmmi5\scriptfont1=\pyxfontsfb\font\pyxfontssfb=cmmi5\scriptscriptfont1=\pyxfontssfb\font\pyxfonttfc=cmsy7\textfont2=\pyxfonttfc\font\pyxfontsfc=cmsy5\scriptfont2=\pyxfontsfc\font\pyxfontssfc=cmsy5\scriptscriptfont2=\pyxfontssfc\font\pyxfonttfd=cmex10\textfont3=\pyxfonttfd\font\pyxfontsfd=cmex10\scriptfont3=\pyxfontsfd\font\pyxfontssfd=cmex10\scriptscriptfont3=\pyxfontssfd}
\def\footnotesize{\font\pyxfont=cmr8\pyxfont\font\pyxfonttfa=cmr8\textfont0=\pyxfonttfa\font\pyxfontsfa=cmr6\scriptfont0=\pyxfontsfa\font\pyxfontssfa=cmr5\scriptscriptfont0=\pyxfontssfa\font\pyxfonttfb=cmmi8\textfont1=\pyxfonttfb\font\pyxfontsfb=cmmi6\scriptfont1=\pyxfontsfb\font\pyxfontssfb=cmmi5\scriptscriptfont1=\pyxfontssfb\font\pyxfonttfc=cmsy8\textfont2=\pyxfonttfc\font\pyxfontsfc=cmsy6\scriptfont2=\pyxfontsfc\font\pyxfontssfc=cmsy5\scriptscriptfont2=\pyxfontssfc\font\pyxfonttfd=cmex10\textfont3=\pyxfonttfd\font\pyxfontsfd=cmex10\scriptfont3=\pyxfontsfd\font\pyxfontssfd=cmex10\scriptscriptfont3=\pyxfontssfd}
\def\small{\font\pyxfont=cmr9\pyxfont\font\pyxfonttfa=cmr9\textfont0=\pyxfonttfa\font\pyxfontsfa=cmr6\scriptfont0=\pyxfontsfa\font\pyxfontssfa=cmr5\scriptscriptfont0=\pyxfontssfa\font\pyxfonttfb=cmmi9\textfont1=\pyxfonttfb\font\pyxfontsfb=cmmi6\scriptfont1=\pyxfontsfb\font\pyxfontssfb=cmmi5\scriptscriptfont1=\pyxfontssfb\font\pyxfonttfc=cmsy9\textfont2=\pyxfonttfc\font\pyxfontsfc=cmsy6\scriptfont2=\pyxfontsfc\font\pyxfontssfc=cmsy5\scriptscriptfont2=\pyxfontssfc\font\pyxfonttfd=cmex10\textfont3=\pyxfonttfd\font\pyxfontsfd=cmex10\scriptfont3=\pyxfontsfd\font\pyxfontssfd=cmex10\scriptscriptfont3=\pyxfontssfd}
\def\normalsize{\font\pyxfont=cmr10\pyxfont\font\pyxfonttfa=cmr10\textfont0=\pyxfonttfa\font\pyxfontsfa=cmr7\scriptfont0=\pyxfontsfa\font\pyxfontssfa=cmr5\scriptscriptfont0=\pyxfontssfa\font\pyxfonttfb=cmmi10\textfont1=\pyxfonttfb\font\pyxfontsfb=cmmi7\scriptfont1=\pyxfontsfb\font\pyxfontssfb=cmmi5\scriptscriptfont1=\pyxfontssfb\font\pyxfonttfc=cmsy10\textfont2=\pyxfonttfc\font\pyxfontsfc=cmsy7\scriptfont2=\pyxfontsfc\font\pyxfontssfc=cmsy5\scriptscriptfont2=\pyxfontssfc\font\pyxfonttfd=cmex10\textfont3=\pyxfonttfd\font\pyxfontsfd=cmex10\scriptfont3=\pyxfontsfd\font\pyxfontssfd=cmex10\scriptscriptfont3=\pyxfontssfd}
\def\large{\font\pyxfont=cmr12\pyxfont\font\pyxfonttfa=cmr12\textfont0=\pyxfonttfa\font\pyxfontsfa=cmr8\scriptfont0=\pyxfontsfa\font\pyxfontssfa=cmr6\scriptscriptfont0=\pyxfontssfa\font\pyxfonttfb=cmmi12\textfont1=\pyxfonttfb\font\pyxfontsfb=cmmi8\scriptfont1=\pyxfontsfb\font\pyxfontssfb=cmmi6\scriptscriptfont1=\pyxfontssfb\font\pyxfonttfc=cmsy10 at 12.0pt\textfont2=\pyxfonttfc\font\pyxfontsfc=cmsy8\scriptfont2=\pyxfontsfc\font\pyxfontssfc=cmsy6\scriptscriptfont2=\pyxfontssfc\font\pyxfonttfd=cmex10\textfont3=\pyxfonttfd\font\pyxfontsfd=cmex10\scriptfont3=\pyxfontsfd\font\pyxfontssfd=cmex10\scriptscriptfont3=\pyxfontssfd}
\def\Large{\font\pyxfont=cmr12 at 14.4pt\pyxfont\font\pyxfonttfa=cmr12 at 14.4pt\textfont0=\pyxfonttfa\font\pyxfontsfa=cmr10\scriptfont0=\pyxfontsfa\font\pyxfontssfa=cmr7\scriptscriptfont0=\pyxfontssfa\font\pyxfonttfb=cmmi12 at 14.4pt\textfont1=\pyxfonttfb\font\pyxfontsfb=cmmi10\scriptfont1=\pyxfontsfb\font\pyxfontssfb=cmmi7\scriptscriptfont1=\pyxfontssfb\font\pyxfonttfc=cmsy10 at 14.4pt\textfont2=\pyxfonttfc\font\pyxfontsfc=cmsy10\scriptfont2=\pyxfontsfc\font\pyxfontssfc=cmsy7\scriptscriptfont2=\pyxfontssfc\font\pyxfonttfd=cmex10\textfont3=\pyxfonttfd\font\pyxfontsfd=cmex10\scriptfont3=\pyxfontsfd\font\pyxfontssfd=cmex10\scriptscriptfont3=\pyxfontssfd}
\def\LARGE{\font\pyxfont=cmr17\pyxfont\font\pyxfonttfa=cmr17\textfont0=\pyxfonttfa\font\pyxfontsfa=cmr12\scriptfont0=\pyxfontsfa\font\pyxfontssfa=cmr10\scriptscriptfont0=\pyxfontssfa\font\pyxfonttfb=cmmi12 at 17.28pt\textfont1=\pyxfonttfb\font\pyxfontsfb=cmmi12\scriptfont1=\pyxfontsfb\font\pyxfontssfb=cmmi10\scriptscriptfont1=\pyxfontssfb\font\pyxfonttfc=cmsy10 at 17.28pt\textfont2=\pyxfonttfc\font\pyxfontsfc=cmsy10 at 12.0pt\scriptfont2=\pyxfontsfc\font\pyxfontssfc=cmsy10\scriptscriptfont2=\pyxfontssfc\font\pyxfonttfd=cmex10\textfont3=\pyxfonttfd\font\pyxfontsfd=cmex10\scriptfont3=\pyxfontsfd\font\pyxfontssfd=cmex10\scriptscriptfont3=\pyxfontssfd}
\def\huge{\font\pyxfont=cmr17 at 20.74pt\pyxfont\font\pyxfonttfa=cmr17 at 20.74pt\textfont0=\pyxfonttfa\font\pyxfontsfa=cmr12 at 14.4pt\scriptfont0=\pyxfontsfa\font\pyxfontssfa=cmr12\scriptscriptfont0=\pyxfontssfa\font\pyxfonttfb=cmmi12 at 20.74pt\textfont1=\pyxfonttfb\font\pyxfontsfb=cmmi12 at 14.4pt\scriptfont1=\pyxfontsfb\font\pyxfontssfb=cmmi12\scriptscriptfont1=\pyxfontssfb\font\pyxfonttfc=cmsy10 at 20.74pt\textfont2=\pyxfonttfc\font\pyxfontsfc=cmsy10 at 14.4pt\scriptfont2=\pyxfontsfc\font\pyxfontssfc=cmsy10 at 12.0pt\scriptscriptfont2=\pyxfontssfc\font\pyxfonttfd=cmex10\textfont3=\pyxfonttfd\font\pyxfontsfd=cmex10\scriptfont3=\pyxfontsfd\font\pyxfontssfd=cmex10\scriptscriptfont3=\pyxfontssfd}
\def\Huge{\font\pyxfont=cmr17 at 24.88pt\pyxfont\font\pyxfonttfa=cmr17 at 24.88pt\textfont0=\pyxfonttfa\font\pyxfontsfa=cmr17 at 20.74pt\scriptfont0=\pyxfontsfa\font\pyxfontssfa=cmr17\scriptscriptfont0=\pyxfontssfa\font\pyxfonttfb=cmmi12 at 24.88pt\textfont1=\pyxfonttfb\font\pyxfontsfb=cmmi12 at 20.74pt\scriptfont1=\pyxfontsfb\font\pyxfontssfb=cmmi12 at 17.28pt\scriptscriptfont1=\pyxfontssfb\font\pyxfonttfc=cmsy10 at 24.88pt\textfont2=\pyxfonttfc\font\pyxfontsfc=cmsy10 at 20.74pt\scriptfont2=\pyxfontsfc\font\pyxfontssfc=cmsy10 at 17.28pt\scriptscriptfont2=\pyxfontssfc\font\pyxfonttfd=cmex10\textfont3=\pyxfonttfd\font\pyxfontsfd=cmex10\scriptfont3=\pyxfontsfd\font\pyxfontssfd=cmex10\scriptscriptfont3=\pyxfontssfd}
%
\PyXInput{2}%
\normalsize%
%
\PyXInput{3}%
\newdimen\linewidth\newdimen\textwidth%
%
\PyXInput{4}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}$\displaystyle{0.1}$}}\lower\PyXDimenVShift\box0%
}{1}%
\PyXInput{5}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}$\displaystyle{1}$}}\lower\PyXDimenVShift\box0%
}{2}%
\PyXInput{6}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}$\displaystyle{10}$}}\lower\PyXDimenVShift\box0%
}{3}%
\PyXInput{7}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}$\displaystyle{100}$}}\lower\PyXDimenVShift\box0%
}{4}%
\PyXInput{8}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}$\displaystyle{1000}$}}\lower\PyXDimenVShift\box0%
}{5}%
\PyXInput{9}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}output power density ($\times10^{10}$) [W/Hz]}}\lower\PyXDimenVShift\box0%
}{6}%
\PyXInput{10}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}$\displaystyle{{0}}$}}\lower\PyXDimenVShift\box0%
}{7}%
\PyXInput{11}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}$\displaystyle{{{2.5}\cdot10^{3}}}$}}\lower\PyXDimenVShift\box0%
}{8}%
\PyXInput{12}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}$\displaystyle{{{5}\cdot10^{3}}}$}}\lower\PyXDimenVShift\box0%
}{9}%
\PyXInput{13}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}$\displaystyle{{{7.5}\cdot10^{3}}}$}}\lower\PyXDimenVShift\box0%
}{10}%
\PyXInput{14}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}$\displaystyle{{{1}\cdot10^{4}}}$}}\lower\PyXDimenVShift\box0%
}{11}%
\PyXInput{15}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}active Q}}\lower\PyXDimenVShift\box0%
}{12}%
\PyXInput{16}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}$\displaystyle{{0}}$}}\lower\PyXDimenVShift\box0%
}{13}%
\PyXInput{17}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}$\displaystyle{{{2}\cdot10^{3}}}$}}\lower\PyXDimenVShift\box0%
}{14}%
\PyXInput{18}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}$\displaystyle{{{4}\cdot10^{3}}}$}}\lower\PyXDimenVShift\box0%
}{15}%
\PyXInput{19}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}$\displaystyle{{{6}\cdot10^{3}}}$}}\lower\PyXDimenVShift\box0%
}{16}%
\PyXInput{20}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}$\displaystyle{{{8}\cdot10^{3}}}$}}\lower\PyXDimenVShift\box0%
}{17}%
\PyXInput{21}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}$\displaystyle{{{1}\cdot10^{4}}}$}}\lower\PyXDimenVShift\box0%
}{18}%
\PyXInput{22}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}active Q}}\lower\PyXDimenVShift\box0%
}{19}%
\PyXInput{23}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{signal}}\lower\PyXDimenVShift\box0%
}{20}%
\PyXInput{24}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{noise}}\lower\PyXDimenVShift\box0%
}{21}%
\PyXInput{25}%
\end%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% preparing restart of tex
% PyX 0.12.1 texdebug file
% mode: tex
% date: Tue May 12 13:48:59 2015
\scrollmode
\raiseerror%
\def\PyX{P\kern-.3em\lower.5ex\hbox{Y}\kern-.18em X}%
\gdef\PyXBoxHAlign{0}%
\newbox\PyXBox%
\newbox\PyXBoxHAligned%
\newdimen\PyXDimenHAlignLT%
\newdimen\PyXDimenHAlignRT%
\gdef\PyXFlushHAlign{0}%
\def\PyXragged{%
\leftskip=0pt plus \PyXFlushHAlign fil%
\rightskip=0pt plus 1fil%
\advance\rightskip0pt plus -\PyXFlushHAlign fil%
\parfillskip=0pt%
\pretolerance=9999%
\tolerance=9999%
\parindent=0pt%
\hyphenpenalty=9999%
\exhyphenpenalty=9999}%
\newbox\PyXBoxVBox%
\newdimen\PyXDimenVBox%
\newbox\PyXBoxVAlign%
\newdimen\PyXDimenVAlign%
\newdimen\PyXDimenVShift%
\long\def\ProcessPyXBox#1#2{%
\setbox\PyXBox=\hbox{{#1}}%
\PyXDimenHAlignLT=\PyXBoxHAlign\wd\PyXBox%
\PyXDimenHAlignRT=\wd\PyXBox%
\advance\PyXDimenHAlignRT by -\PyXDimenHAlignLT%
\gdef\PyXBoxHAlign{0}%
\immediate\write16{PyXBox:page=#2,lt=\the\PyXDimenHAlignLT,rt=\the\PyXDimenHAlignRT,ht=\the\ht\PyXBox,dp=\the\dp\PyXBox:}%
\setbox\PyXBoxHAligned=\hbox{\kern-\PyXDimenHAlignLT\box\PyXBox}%
\ht\PyXBoxHAligned0pt%
{\count0=80\count1=121\count2=88\count3=#2\shipout\box\PyXBoxHAligned}}%
\def\PyXInput#1{\immediate\write16{PyXInputMarker:executeid=#1:}}%
\def\PyXMarker#1{\hskip0pt\special{PyX:marker #1}}%%
\PyXInput{1}%
% This automatically generated file is part of PyX (http://pyx.sourceforge.net/).
% \latexstylename="10pt"
% \latexclassname="article"
% \latexclassopt="10pt"
% \latexinit=""
\def\tiny{\font\pyxfont=cmr5\pyxfont\font\pyxfonttfa=cmr5\textfont0=\pyxfonttfa\font\pyxfontsfa=cmr5\scriptfont0=\pyxfontsfa\font\pyxfontssfa=cmr5\scriptscriptfont0=\pyxfontssfa\font\pyxfonttfb=cmmi5\textfont1=\pyxfonttfb\font\pyxfontsfb=cmmi5\scriptfont1=\pyxfontsfb\font\pyxfontssfb=cmmi5\scriptscriptfont1=\pyxfontssfb\font\pyxfonttfc=cmsy5\textfont2=\pyxfonttfc\font\pyxfontsfc=cmsy5\scriptfont2=\pyxfontsfc\font\pyxfontssfc=cmsy5\scriptscriptfont2=\pyxfontssfc\font\pyxfonttfd=cmex10\textfont3=\pyxfonttfd\font\pyxfontsfd=cmex10\scriptfont3=\pyxfontsfd\font\pyxfontssfd=cmex10\scriptscriptfont3=\pyxfontssfd}
\def\scriptsize{\font\pyxfont=cmr7\pyxfont\font\pyxfonttfa=cmr7\textfont0=\pyxfonttfa\font\pyxfontsfa=cmr5\scriptfont0=\pyxfontsfa\font\pyxfontssfa=cmr5\scriptscriptfont0=\pyxfontssfa\font\pyxfonttfb=cmmi7\textfont1=\pyxfonttfb\font\pyxfontsfb=cmmi5\scriptfont1=\pyxfontsfb\font\pyxfontssfb=cmmi5\scriptscriptfont1=\pyxfontssfb\font\pyxfonttfc=cmsy7\textfont2=\pyxfonttfc\font\pyxfontsfc=cmsy5\scriptfont2=\pyxfontsfc\font\pyxfontssfc=cmsy5\scriptscriptfont2=\pyxfontssfc\font\pyxfonttfd=cmex10\textfont3=\pyxfonttfd\font\pyxfontsfd=cmex10\scriptfont3=\pyxfontsfd\font\pyxfontssfd=cmex10\scriptscriptfont3=\pyxfontssfd}
\def\footnotesize{\font\pyxfont=cmr8\pyxfont\font\pyxfonttfa=cmr8\textfont0=\pyxfonttfa\font\pyxfontsfa=cmr6\scriptfont0=\pyxfontsfa\font\pyxfontssfa=cmr5\scriptscriptfont0=\pyxfontssfa\font\pyxfonttfb=cmmi8\textfont1=\pyxfonttfb\font\pyxfontsfb=cmmi6\scriptfont1=\pyxfontsfb\font\pyxfontssfb=cmmi5\scriptscriptfont1=\pyxfontssfb\font\pyxfonttfc=cmsy8\textfont2=\pyxfonttfc\font\pyxfontsfc=cmsy6\scriptfont2=\pyxfontsfc\font\pyxfontssfc=cmsy5\scriptscriptfont2=\pyxfontssfc\font\pyxfonttfd=cmex10\textfont3=\pyxfonttfd\font\pyxfontsfd=cmex10\scriptfont3=\pyxfontsfd\font\pyxfontssfd=cmex10\scriptscriptfont3=\pyxfontssfd}
\def\small{\font\pyxfont=cmr9\pyxfont\font\pyxfonttfa=cmr9\textfont0=\pyxfonttfa\font\pyxfontsfa=cmr6\scriptfont0=\pyxfontsfa\font\pyxfontssfa=cmr5\scriptscriptfont0=\pyxfontssfa\font\pyxfonttfb=cmmi9\textfont1=\pyxfonttfb\font\pyxfontsfb=cmmi6\scriptfont1=\pyxfontsfb\font\pyxfontssfb=cmmi5\scriptscriptfont1=\pyxfontssfb\font\pyxfonttfc=cmsy9\textfont2=\pyxfonttfc\font\pyxfontsfc=cmsy6\scriptfont2=\pyxfontsfc\font\pyxfontssfc=cmsy5\scriptscriptfont2=\pyxfontssfc\font\pyxfonttfd=cmex10\textfont3=\pyxfonttfd\font\pyxfontsfd=cmex10\scriptfont3=\pyxfontsfd\font\pyxfontssfd=cmex10\scriptscriptfont3=\pyxfontssfd}
\def\normalsize{\font\pyxfont=cmr10\pyxfont\font\pyxfonttfa=cmr10\textfont0=\pyxfonttfa\font\pyxfontsfa=cmr7\scriptfont0=\pyxfontsfa\font\pyxfontssfa=cmr5\scriptscriptfont0=\pyxfontssfa\font\pyxfonttfb=cmmi10\textfont1=\pyxfonttfb\font\pyxfontsfb=cmmi7\scriptfont1=\pyxfontsfb\font\pyxfontssfb=cmmi5\scriptscriptfont1=\pyxfontssfb\font\pyxfonttfc=cmsy10\textfont2=\pyxfonttfc\font\pyxfontsfc=cmsy7\scriptfont2=\pyxfontsfc\font\pyxfontssfc=cmsy5\scriptscriptfont2=\pyxfontssfc\font\pyxfonttfd=cmex10\textfont3=\pyxfonttfd\font\pyxfontsfd=cmex10\scriptfont3=\pyxfontsfd\font\pyxfontssfd=cmex10\scriptscriptfont3=\pyxfontssfd}
\def\large{\font\pyxfont=cmr12\pyxfont\font\pyxfonttfa=cmr12\textfont0=\pyxfonttfa\font\pyxfontsfa=cmr8\scriptfont0=\pyxfontsfa\font\pyxfontssfa=cmr6\scriptscriptfont0=\pyxfontssfa\font\pyxfonttfb=cmmi12\textfont1=\pyxfonttfb\font\pyxfontsfb=cmmi8\scriptfont1=\pyxfontsfb\font\pyxfontssfb=cmmi6\scriptscriptfont1=\pyxfontssfb\font\pyxfonttfc=cmsy10 at 12.0pt\textfont2=\pyxfonttfc\font\pyxfontsfc=cmsy8\scriptfont2=\pyxfontsfc\font\pyxfontssfc=cmsy6\scriptscriptfont2=\pyxfontssfc\font\pyxfonttfd=cmex10\textfont3=\pyxfonttfd\font\pyxfontsfd=cmex10\scriptfont3=\pyxfontsfd\font\pyxfontssfd=cmex10\scriptscriptfont3=\pyxfontssfd}
\def\Large{\font\pyxfont=cmr12 at 14.4pt\pyxfont\font\pyxfonttfa=cmr12 at 14.4pt\textfont0=\pyxfonttfa\font\pyxfontsfa=cmr10\scriptfont0=\pyxfontsfa\font\pyxfontssfa=cmr7\scriptscriptfont0=\pyxfontssfa\font\pyxfonttfb=cmmi12 at 14.4pt\textfont1=\pyxfonttfb\font\pyxfontsfb=cmmi10\scriptfont1=\pyxfontsfb\font\pyxfontssfb=cmmi7\scriptscriptfont1=\pyxfontssfb\font\pyxfonttfc=cmsy10 at 14.4pt\textfont2=\pyxfonttfc\font\pyxfontsfc=cmsy10\scriptfont2=\pyxfontsfc\font\pyxfontssfc=cmsy7\scriptscriptfont2=\pyxfontssfc\font\pyxfonttfd=cmex10\textfont3=\pyxfonttfd\font\pyxfontsfd=cmex10\scriptfont3=\pyxfontsfd\font\pyxfontssfd=cmex10\scriptscriptfont3=\pyxfontssfd}
\def\LARGE{\font\pyxfont=cmr17\pyxfont\font\pyxfonttfa=cmr17\textfont0=\pyxfonttfa\font\pyxfontsfa=cmr12\scriptfont0=\pyxfontsfa\font\pyxfontssfa=cmr10\scriptscriptfont0=\pyxfontssfa\font\pyxfonttfb=cmmi12 at 17.28pt\textfont1=\pyxfonttfb\font\pyxfontsfb=cmmi12\scriptfont1=\pyxfontsfb\font\pyxfontssfb=cmmi10\scriptscriptfont1=\pyxfontssfb\font\pyxfonttfc=cmsy10 at 17.28pt\textfont2=\pyxfonttfc\font\pyxfontsfc=cmsy10 at 12.0pt\scriptfont2=\pyxfontsfc\font\pyxfontssfc=cmsy10\scriptscriptfont2=\pyxfontssfc\font\pyxfonttfd=cmex10\textfont3=\pyxfonttfd\font\pyxfontsfd=cmex10\scriptfont3=\pyxfontsfd\font\pyxfontssfd=cmex10\scriptscriptfont3=\pyxfontssfd}
\def\huge{\font\pyxfont=cmr17 at 20.74pt\pyxfont\font\pyxfonttfa=cmr17 at 20.74pt\textfont0=\pyxfonttfa\font\pyxfontsfa=cmr12 at 14.4pt\scriptfont0=\pyxfontsfa\font\pyxfontssfa=cmr12\scriptscriptfont0=\pyxfontssfa\font\pyxfonttfb=cmmi12 at 20.74pt\textfont1=\pyxfonttfb\font\pyxfontsfb=cmmi12 at 14.4pt\scriptfont1=\pyxfontsfb\font\pyxfontssfb=cmmi12\scriptscriptfont1=\pyxfontssfb\font\pyxfonttfc=cmsy10 at 20.74pt\textfont2=\pyxfonttfc\font\pyxfontsfc=cmsy10 at 14.4pt\scriptfont2=\pyxfontsfc\font\pyxfontssfc=cmsy10 at 12.0pt\scriptscriptfont2=\pyxfontssfc\font\pyxfonttfd=cmex10\textfont3=\pyxfonttfd\font\pyxfontsfd=cmex10\scriptfont3=\pyxfontsfd\font\pyxfontssfd=cmex10\scriptscriptfont3=\pyxfontssfd}
\def\Huge{\font\pyxfont=cmr17 at 24.88pt\pyxfont\font\pyxfonttfa=cmr17 at 24.88pt\textfont0=\pyxfonttfa\font\pyxfontsfa=cmr17 at 20.74pt\scriptfont0=\pyxfontsfa\font\pyxfontssfa=cmr17\scriptscriptfont0=\pyxfontssfa\font\pyxfonttfb=cmmi12 at 24.88pt\textfont1=\pyxfonttfb\font\pyxfontsfb=cmmi12 at 20.74pt\scriptfont1=\pyxfontsfb\font\pyxfontssfb=cmmi12 at 17.28pt\scriptscriptfont1=\pyxfontssfb\font\pyxfonttfc=cmsy10 at 24.88pt\textfont2=\pyxfonttfc\font\pyxfontsfc=cmsy10 at 20.74pt\scriptfont2=\pyxfontsfc\font\pyxfontssfc=cmsy10 at 17.28pt\scriptscriptfont2=\pyxfontssfc\font\pyxfonttfd=cmex10\textfont3=\pyxfonttfd\font\pyxfontsfd=cmex10\scriptfont3=\pyxfontsfd\font\pyxfontssfd=cmex10\scriptscriptfont3=\pyxfontssfd}
%
\PyXInput{2}%
\normalsize%
%
\PyXInput{3}%
\newdimen\linewidth\newdimen\textwidth%
%
\PyXInput{4}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}$\displaystyle{1}$}}\lower\PyXDimenVShift\box0%
}{1}%
\PyXInput{5}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}$\displaystyle{2}$}}\lower\PyXDimenVShift\box0%
}{2}%
\PyXInput{6}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}$\displaystyle{5}$}}\lower\PyXDimenVShift\box0%
}{3}%
\PyXInput{7}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}$\displaystyle{10}$}}\lower\PyXDimenVShift\box0%
}{4}%
\PyXInput{8}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}$\displaystyle{20}$}}\lower\PyXDimenVShift\box0%
}{5}%
\PyXInput{9}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}$\displaystyle{50}$}}\lower\PyXDimenVShift\box0%
}{6}%
\PyXInput{10}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}$\displaystyle{100}$}}\lower\PyXDimenVShift\box0%
}{7}%
\PyXInput{11}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}output power density ($\times10^{10}$) [W/Hz]}}\lower\PyXDimenVShift\box0%
}{8}%
\PyXInput{12}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}$\displaystyle{1}$}}\lower\PyXDimenVShift\box0%
}{9}%
\PyXInput{13}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}$\displaystyle{10}$}}\lower\PyXDimenVShift\box0%
}{10}%
\PyXInput{14}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}$\displaystyle{100}$}}\lower\PyXDimenVShift\box0%
}{11}%
\PyXInput{15}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}output power density ($\times10^{10}$) [W/Hz]}}\lower\PyXDimenVShift\box0%
}{12}%
\PyXInput{16}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}$\displaystyle{1000}$}}\lower\PyXDimenVShift\box0%
}{13}%
\PyXInput{17}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}$\displaystyle{2000}$}}\lower\PyXDimenVShift\box0%
}{14}%
\PyXInput{18}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}$\displaystyle{3000}$}}\lower\PyXDimenVShift\box0%
}{15}%
\PyXInput{19}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}$\displaystyle{4000}$}}\lower\PyXDimenVShift\box0%
}{16}%
\PyXInput{20}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}$\displaystyle{5000}$}}\lower\PyXDimenVShift\box0%
}{17}%
\PyXInput{21}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}$\displaystyle{6000}$}}\lower\PyXDimenVShift\box0%
}{18}%
\PyXInput{22}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}active Q}}\lower\PyXDimenVShift\box0%
}{19}%
\PyXInput{23}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{signal}}\lower\PyXDimenVShift\box0%
}{20}%
\PyXInput{24}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{noise}}\lower\PyXDimenVShift\box0%
}{21}%
\PyXInput{25}%
\end%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% preparing restart of tex
% PyX 0.12.1 texdebug file
% mode: tex
% date: Tue May 12 13:49:16 2015
\scrollmode
\raiseerror%
\def\PyX{P\kern-.3em\lower.5ex\hbox{Y}\kern-.18em X}%
\gdef\PyXBoxHAlign{0}%
\newbox\PyXBox%
\newbox\PyXBoxHAligned%
\newdimen\PyXDimenHAlignLT%
\newdimen\PyXDimenHAlignRT%
\gdef\PyXFlushHAlign{0}%
\def\PyXragged{%
\leftskip=0pt plus \PyXFlushHAlign fil%
\rightskip=0pt plus 1fil%
\advance\rightskip0pt plus -\PyXFlushHAlign fil%
\parfillskip=0pt%
\pretolerance=9999%
\tolerance=9999%
\parindent=0pt%
\hyphenpenalty=9999%
\exhyphenpenalty=9999}%
\newbox\PyXBoxVBox%
\newdimen\PyXDimenVBox%
\newbox\PyXBoxVAlign%
\newdimen\PyXDimenVAlign%
\newdimen\PyXDimenVShift%
\long\def\ProcessPyXBox#1#2{%
\setbox\PyXBox=\hbox{{#1}}%
\PyXDimenHAlignLT=\PyXBoxHAlign\wd\PyXBox%
\PyXDimenHAlignRT=\wd\PyXBox%
\advance\PyXDimenHAlignRT by -\PyXDimenHAlignLT%
\gdef\PyXBoxHAlign{0}%
\immediate\write16{PyXBox:page=#2,lt=\the\PyXDimenHAlignLT,rt=\the\PyXDimenHAlignRT,ht=\the\ht\PyXBox,dp=\the\dp\PyXBox:}%
\setbox\PyXBoxHAligned=\hbox{\kern-\PyXDimenHAlignLT\box\PyXBox}%
\ht\PyXBoxHAligned0pt%
{\count0=80\count1=121\count2=88\count3=#2\shipout\box\PyXBoxHAligned}}%
\def\PyXInput#1{\immediate\write16{PyXInputMarker:executeid=#1:}}%
\def\PyXMarker#1{\hskip0pt\special{PyX:marker #1}}%%
\PyXInput{1}%
% This automatically generated file is part of PyX (http://pyx.sourceforge.net/).
% \latexstylename="10pt"
% \latexclassname="article"
% \latexclassopt="10pt"
% \latexinit=""
\def\tiny{\font\pyxfont=cmr5\pyxfont\font\pyxfonttfa=cmr5\textfont0=\pyxfonttfa\font\pyxfontsfa=cmr5\scriptfont0=\pyxfontsfa\font\pyxfontssfa=cmr5\scriptscriptfont0=\pyxfontssfa\font\pyxfonttfb=cmmi5\textfont1=\pyxfonttfb\font\pyxfontsfb=cmmi5\scriptfont1=\pyxfontsfb\font\pyxfontssfb=cmmi5\scriptscriptfont1=\pyxfontssfb\font\pyxfonttfc=cmsy5\textfont2=\pyxfonttfc\font\pyxfontsfc=cmsy5\scriptfont2=\pyxfontsfc\font\pyxfontssfc=cmsy5\scriptscriptfont2=\pyxfontssfc\font\pyxfonttfd=cmex10\textfont3=\pyxfonttfd\font\pyxfontsfd=cmex10\scriptfont3=\pyxfontsfd\font\pyxfontssfd=cmex10\scriptscriptfont3=\pyxfontssfd}
\def\scriptsize{\font\pyxfont=cmr7\pyxfont\font\pyxfonttfa=cmr7\textfont0=\pyxfonttfa\font\pyxfontsfa=cmr5\scriptfont0=\pyxfontsfa\font\pyxfontssfa=cmr5\scriptscriptfont0=\pyxfontssfa\font\pyxfonttfb=cmmi7\textfont1=\pyxfonttfb\font\pyxfontsfb=cmmi5\scriptfont1=\pyxfontsfb\font\pyxfontssfb=cmmi5\scriptscriptfont1=\pyxfontssfb\font\pyxfonttfc=cmsy7\textfont2=\pyxfonttfc\font\pyxfontsfc=cmsy5\scriptfont2=\pyxfontsfc\font\pyxfontssfc=cmsy5\scriptscriptfont2=\pyxfontssfc\font\pyxfonttfd=cmex10\textfont3=\pyxfonttfd\font\pyxfontsfd=cmex10\scriptfont3=\pyxfontsfd\font\pyxfontssfd=cmex10\scriptscriptfont3=\pyxfontssfd}
\def\footnotesize{\font\pyxfont=cmr8\pyxfont\font\pyxfonttfa=cmr8\textfont0=\pyxfonttfa\font\pyxfontsfa=cmr6\scriptfont0=\pyxfontsfa\font\pyxfontssfa=cmr5\scriptscriptfont0=\pyxfontssfa\font\pyxfonttfb=cmmi8\textfont1=\pyxfonttfb\font\pyxfontsfb=cmmi6\scriptfont1=\pyxfontsfb\font\pyxfontssfb=cmmi5\scriptscriptfont1=\pyxfontssfb\font\pyxfonttfc=cmsy8\textfont2=\pyxfonttfc\font\pyxfontsfc=cmsy6\scriptfont2=\pyxfontsfc\font\pyxfontssfc=cmsy5\scriptscriptfont2=\pyxfontssfc\font\pyxfonttfd=cmex10\textfont3=\pyxfonttfd\font\pyxfontsfd=cmex10\scriptfont3=\pyxfontsfd\font\pyxfontssfd=cmex10\scriptscriptfont3=\pyxfontssfd}
\def\small{\font\pyxfont=cmr9\pyxfont\font\pyxfonttfa=cmr9\textfont0=\pyxfonttfa\font\pyxfontsfa=cmr6\scriptfont0=\pyxfontsfa\font\pyxfontssfa=cmr5\scriptscriptfont0=\pyxfontssfa\font\pyxfonttfb=cmmi9\textfont1=\pyxfonttfb\font\pyxfontsfb=cmmi6\scriptfont1=\pyxfontsfb\font\pyxfontssfb=cmmi5\scriptscriptfont1=\pyxfontssfb\font\pyxfonttfc=cmsy9\textfont2=\pyxfonttfc\font\pyxfontsfc=cmsy6\scriptfont2=\pyxfontsfc\font\pyxfontssfc=cmsy5\scriptscriptfont2=\pyxfontssfc\font\pyxfonttfd=cmex10\textfont3=\pyxfonttfd\font\pyxfontsfd=cmex10\scriptfont3=\pyxfontsfd\font\pyxfontssfd=cmex10\scriptscriptfont3=\pyxfontssfd}
\def\normalsize{\font\pyxfont=cmr10\pyxfont\font\pyxfonttfa=cmr10\textfont0=\pyxfonttfa\font\pyxfontsfa=cmr7\scriptfont0=\pyxfontsfa\font\pyxfontssfa=cmr5\scriptscriptfont0=\pyxfontssfa\font\pyxfonttfb=cmmi10\textfont1=\pyxfonttfb\font\pyxfontsfb=cmmi7\scriptfont1=\pyxfontsfb\font\pyxfontssfb=cmmi5\scriptscriptfont1=\pyxfontssfb\font\pyxfonttfc=cmsy10\textfont2=\pyxfonttfc\font\pyxfontsfc=cmsy7\scriptfont2=\pyxfontsfc\font\pyxfontssfc=cmsy5\scriptscriptfont2=\pyxfontssfc\font\pyxfonttfd=cmex10\textfont3=\pyxfonttfd\font\pyxfontsfd=cmex10\scriptfont3=\pyxfontsfd\font\pyxfontssfd=cmex10\scriptscriptfont3=\pyxfontssfd}
\def\large{\font\pyxfont=cmr12\pyxfont\font\pyxfonttfa=cmr12\textfont0=\pyxfonttfa\font\pyxfontsfa=cmr8\scriptfont0=\pyxfontsfa\font\pyxfontssfa=cmr6\scriptscriptfont0=\pyxfontssfa\font\pyxfonttfb=cmmi12\textfont1=\pyxfonttfb\font\pyxfontsfb=cmmi8\scriptfont1=\pyxfontsfb\font\pyxfontssfb=cmmi6\scriptscriptfont1=\pyxfontssfb\font\pyxfonttfc=cmsy10 at 12.0pt\textfont2=\pyxfonttfc\font\pyxfontsfc=cmsy8\scriptfont2=\pyxfontsfc\font\pyxfontssfc=cmsy6\scriptscriptfont2=\pyxfontssfc\font\pyxfonttfd=cmex10\textfont3=\pyxfonttfd\font\pyxfontsfd=cmex10\scriptfont3=\pyxfontsfd\font\pyxfontssfd=cmex10\scriptscriptfont3=\pyxfontssfd}
\def\Large{\font\pyxfont=cmr12 at 14.4pt\pyxfont\font\pyxfonttfa=cmr12 at 14.4pt\textfont0=\pyxfonttfa\font\pyxfontsfa=cmr10\scriptfont0=\pyxfontsfa\font\pyxfontssfa=cmr7\scriptscriptfont0=\pyxfontssfa\font\pyxfonttfb=cmmi12 at 14.4pt\textfont1=\pyxfonttfb\font\pyxfontsfb=cmmi10\scriptfont1=\pyxfontsfb\font\pyxfontssfb=cmmi7\scriptscriptfont1=\pyxfontssfb\font\pyxfonttfc=cmsy10 at 14.4pt\textfont2=\pyxfonttfc\font\pyxfontsfc=cmsy10\scriptfont2=\pyxfontsfc\font\pyxfontssfc=cmsy7\scriptscriptfont2=\pyxfontssfc\font\pyxfonttfd=cmex10\textfont3=\pyxfonttfd\font\pyxfontsfd=cmex10\scriptfont3=\pyxfontsfd\font\pyxfontssfd=cmex10\scriptscriptfont3=\pyxfontssfd}
\def\LARGE{\font\pyxfont=cmr17\pyxfont\font\pyxfonttfa=cmr17\textfont0=\pyxfonttfa\font\pyxfontsfa=cmr12\scriptfont0=\pyxfontsfa\font\pyxfontssfa=cmr10\scriptscriptfont0=\pyxfontssfa\font\pyxfonttfb=cmmi12 at 17.28pt\textfont1=\pyxfonttfb\font\pyxfontsfb=cmmi12\scriptfont1=\pyxfontsfb\font\pyxfontssfb=cmmi10\scriptscriptfont1=\pyxfontssfb\font\pyxfonttfc=cmsy10 at 17.28pt\textfont2=\pyxfonttfc\font\pyxfontsfc=cmsy10 at 12.0pt\scriptfont2=\pyxfontsfc\font\pyxfontssfc=cmsy10\scriptscriptfont2=\pyxfontssfc\font\pyxfonttfd=cmex10\textfont3=\pyxfonttfd\font\pyxfontsfd=cmex10\scriptfont3=\pyxfontsfd\font\pyxfontssfd=cmex10\scriptscriptfont3=\pyxfontssfd}
\def\huge{\font\pyxfont=cmr17 at 20.74pt\pyxfont\font\pyxfonttfa=cmr17 at 20.74pt\textfont0=\pyxfonttfa\font\pyxfontsfa=cmr12 at 14.4pt\scriptfont0=\pyxfontsfa\font\pyxfontssfa=cmr12\scriptscriptfont0=\pyxfontssfa\font\pyxfonttfb=cmmi12 at 20.74pt\textfont1=\pyxfonttfb\font\pyxfontsfb=cmmi12 at 14.4pt\scriptfont1=\pyxfontsfb\font\pyxfontssfb=cmmi12\scriptscriptfont1=\pyxfontssfb\font\pyxfonttfc=cmsy10 at 20.74pt\textfont2=\pyxfonttfc\font\pyxfontsfc=cmsy10 at 14.4pt\scriptfont2=\pyxfontsfc\font\pyxfontssfc=cmsy10 at 12.0pt\scriptscriptfont2=\pyxfontssfc\font\pyxfonttfd=cmex10\textfont3=\pyxfonttfd\font\pyxfontsfd=cmex10\scriptfont3=\pyxfontsfd\font\pyxfontssfd=cmex10\scriptscriptfont3=\pyxfontssfd}
\def\Huge{\font\pyxfont=cmr17 at 24.88pt\pyxfont\font\pyxfonttfa=cmr17 at 24.88pt\textfont0=\pyxfonttfa\font\pyxfontsfa=cmr17 at 20.74pt\scriptfont0=\pyxfontsfa\font\pyxfontssfa=cmr17\scriptscriptfont0=\pyxfontssfa\font\pyxfonttfb=cmmi12 at 24.88pt\textfont1=\pyxfonttfb\font\pyxfontsfb=cmmi12 at 20.74pt\scriptfont1=\pyxfontsfb\font\pyxfontssfb=cmmi12 at 17.28pt\scriptscriptfont1=\pyxfontssfb\font\pyxfonttfc=cmsy10 at 24.88pt\textfont2=\pyxfonttfc\font\pyxfontsfc=cmsy10 at 20.74pt\scriptfont2=\pyxfontsfc\font\pyxfontssfc=cmsy10 at 17.28pt\scriptscriptfont2=\pyxfontssfc\font\pyxfonttfd=cmex10\textfont3=\pyxfonttfd\font\pyxfontsfd=cmex10\scriptfont3=\pyxfontsfd\font\pyxfontssfd=cmex10\scriptscriptfont3=\pyxfontssfd}
%
\PyXInput{2}%
\normalsize%
%
\PyXInput{3}%
\newdimen\linewidth\newdimen\textwidth%
%
\PyXInput{4}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}$\displaystyle{1}$}}\lower\PyXDimenVShift\box0%
}{1}%
\PyXInput{5}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}$\displaystyle{2}$}}\lower\PyXDimenVShift\box0%
}{2}%
\PyXInput{6}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}$\displaystyle{5}$}}\lower\PyXDimenVShift\box0%
}{3}%
\PyXInput{7}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}$\displaystyle{10}$}}\lower\PyXDimenVShift\box0%
}{4}%
\PyXInput{8}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}$\displaystyle{20}$}}\lower\PyXDimenVShift\box0%
}{5}%
\PyXInput{9}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}$\displaystyle{50}$}}\lower\PyXDimenVShift\box0%
}{6}%
\PyXInput{10}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}$\displaystyle{100}$}}\lower\PyXDimenVShift\box0%
}{7}%
\PyXInput{11}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}output power density ($\times10^{10}$) [W/Hz]}}\lower\PyXDimenVShift\box0%
}{8}%
\PyXInput{12}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}$\displaystyle{1}$}}\lower\PyXDimenVShift\box0%
}{9}%
\PyXInput{13}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}$\displaystyle{10}$}}\lower\PyXDimenVShift\box0%
}{10}%
\PyXInput{14}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}$\displaystyle{100}$}}\lower\PyXDimenVShift\box0%
}{11}%
\PyXInput{15}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}output power density ($\times10^{10}$) [W/Hz]}}\lower\PyXDimenVShift\box0%
}{12}%
\PyXInput{16}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}$\displaystyle{1000}$}}\lower\PyXDimenVShift\box0%
}{13}%
\PyXInput{17}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}$\displaystyle{2000}$}}\lower\PyXDimenVShift\box0%
}{14}%
\PyXInput{18}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}$\displaystyle{3000}$}}\lower\PyXDimenVShift\box0%
}{15}%
\PyXInput{19}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}$\displaystyle{4000}$}}\lower\PyXDimenVShift\box0%
}{16}%
\PyXInput{20}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}$\displaystyle{5000}$}}\lower\PyXDimenVShift\box0%
}{17}%
\PyXInput{21}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}$\displaystyle{6000}$}}\lower\PyXDimenVShift\box0%
}{18}%
\PyXInput{22}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{\gdef\PyXBoxHAlign{0.50000}\gdef\PyXFlushHAlign{0.50000}\PyXragged{}active Q}}\lower\PyXDimenVShift\box0%
}{19}%
\PyXInput{23}%
\ProcessPyXBox{\setbox0\hbox{$\vcenter{\vrule width0pt}$}\PyXDimenVShift=\ht0\setbox0\hbox{{signal}}\lower\PyXDimenVShift\box0%
}{20}%
\PyXInput{24}%
\ProcessPyXBox{\setbox0\hbox{$\